---
title: "Development"
author: "Toni Price"
date: last-modified
date-format: "[Last updated:] DD MMMM, YYYY"
format: 
  html
vignette: >
  %\VignetteIndexEntry{Development}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

~~~{=comment}
```{r setup}
library(spf)
```
~~~

## Setup

### devtools

1.  For development work, ensure `devtools` is installed. <br>
    From the `R` console:

    ```
    > install.packages("devtools")
    ```

2.  Set up the environment
    - TODO: Implement renv? (Or add install.packages(c("")) instructions)
    - TODO: Add config instructions

### Some useful dev commands

These commands should be executed from the `R` console.

(Note that when `devtools` is installed, the following commands are usually
bound to keyboard shortcuts in RStudio; these are shown in the comments where
applicable)

-  Load the development version of the package:

    ```
    # Ctrl/Cmd + Shift + L
    > devtools::load_all(".")
    ```

-  Run the automated test suite:

    ```
    # Ctrl/Cmd + Shift + T
    > devtools::test(".")
    ```

-  Create package documentation:

    ```
    # Ctrl/Cmd + Shift + D
    > devtools::document(roclets = c('rd', 'collate', 'namespace'))
    ```

-  Build the package vignettes:

    ```
    > devtools::build_vignettes()
    ```

-  Run a package check:

    ```
    # Ctrl/Cmd + Shift + E
    > devtools::check()
    ```
