---
title: "spf"
author: "Toni Price"
date: last-modified
date-format: "[Last updated:] DD MMMM, YYYY"
format: 
  html:
    mermaid: 
      theme: default
vignette: >
  %\VignetteIndexEntry{spf}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

~~~{=comment}
```{r setup}
library(spf)
```
~~~

## Data workflow

### Data phases

```{mermaid}
flowchart LR
  A("Raw") --> B("Wrangled")
  B --> C("Output")
```

1. Initial input is raw data (e.g. '.csv' files, '.nc' files).
2. Raw data is then processed by transforming, merging, etc. (which is
collectively called 'wrangling').
3. The wrangled data is finally used for producing output.

### Environments

For each of these data wrangling phases, there are three possible environments:

- `ref`: Reference data
  - historical data
  - for comparison with data being produced from current code
  - to ensure the converted code is working as expected

- `act`: Actual data
  - data currently being produced by the code
  - can be compared with reference data in automated tests

- `test`: Test data
  - small enough to use for unit/integration tests
  - can be reliably reproduced
